Changes in Patterns of Geometric Sign Groupings as a Model of Social Dynamics Across Aurignacian Phases

### 2025-04-29 endofquarterpaper.qmd
stash
minor edits

trans_network <- function(artifact_data){
    jac <- vegdist(artifact_data %>%  mutate_all(~ as.numeric(. > 0)), "jaccard", binary = TRUE)
    dm <- as.matrix(jac)
    disim <- 1 - dm
    disim[disim<0.1] <- 0
    disim_net <- network(disim,
                         directed=F,
                         ignore.eval=F,
                         names.eval='weight')
    disim_net %v% 'vertex.names' <- row.names(artifact_data)
    set.seed(500)
    final_plot <- ggraph(disim_net, layout="fr") +
      geom_edge_link(aes(width = weight, 
                         alpha=weight), 
                       edge_colour = "black", 
                       show.legend=F) +
      scale_edge_width(range=c(0.5,1.5)) +
      scale_edge_colour_gradient(low = "#CCCCCC",
                                   high = "#000000") +
      geom_node_point(alpha=1, size=5) +
      geom_node_text(aes(label=row.names(artifact_data)), 
                     size=3, 
                     family= "Arial", 
                     repel=T, max.overlaps = 100) +
      theme(text= element_text(size = 15), 
            panel.background = element_rect(fill = 'white')) +
      border(color = "black", size = 0.5)
    return(final_plot)
    }
```

@@

```{r fig.height= 5.5, fig.width= 8}
#| label: fig-neighbor-net
#| fig-cap: "Network analysis of the data"
network_trans <- trans_network(trans_artifact_data)
network_proto <- produce_clusters(proto_artifact_data, proto_unique_data, method = "network")
network_early <- produce_clusters(early_artifact_data, early_unique_data, method = "network")
network_evolved <- produce_clusters(evolved_artifact_data, evolved_unique_data, method = "network")

@@

## Cultural  Strength

@@

The perMANOVA R^2^ statistic shows the proportion of sign variation that can be explained by group membership. The Transitional period has the highest score of `r trans_perm_r` while the Early Aurignacian has the lowest score of `r early_perm_r`. We see a large decline in scores between the Transitional and the Proto-Aurignacian periods, before a slight increase in between the Early and Evolved Aurignacian periods. The pseudo-F statistic quantifies the difference between groups, with the Transitional phase having the highest value of `r trans_perm_f`, and the Early Aurignacian having the lowest value of `r early_perm_f`. Similar to the R^2^ statistic, we see a decline in values between the Transitional and the Proto-Aurignacian periods, before a slight increase in between the Early and Evolved Aurignacian period. The perMANOVA p-value indicates a statistically significant difference in the distribution of signs between the restricted and broad groups for the Proto- , Early, and Evolved Aurignacian phases, but not the Transitional phase (p = `r trans_perm_p`).

@@

## Grouping Dynamics Across the Four Phases

@@

## Dynamics of Social Organization

@@

<-- cite signbase 2.0 -->.

@@

One limitation is the potential ambiguity that arises with some of the sign types. While @Von_Petzinger_2009 argues in favor of Aurignacian geometric signs as symbolic expression, potential ambiguity still remains over whether function is consistent across all sign type occurrences, particularly in regards to the simpler sign types composed from fewer elements. We included the more ambiguous sign types because similar sign occurrence may still serve as evidence of cultural connectivity regardless of function, and additionally, greater quantities of included sign types serves to establish stronger connections [@Gravel-Miguel_Coward_2023]. This ambiguity may be addressed in the future through the usage of the data recently released in SignBase 2.0 covering sites in Germany, which contains more detailed information on the specific order and patterns in which signs occurred on individual objects

\<--

cite signbase 2.0–\>.

>>>>

@@

```{r fig.height= 5.5, fig.width= 8}
#| label: fig-neighbor-net
#| fig-cap: "Network analysis of the data"
network_trans <- produce_clusters(trans_artifact_data, trans_unique_data, method = "network")
network_proto <- produce_clusters(proto_artifact_data, proto_unique_data, method = "network")
network_early <- produce_clusters(early_artifact_data, early_unique_data, method = "network")
network_evolved <- produce_clusters(evolved_artifact_data, evolved_unique_data, method = "network")

@@

## Cultural Connection Strength

@@

The perMANOVA R^2^ statistic shows the proportion of sign variation that can be explained by group membership. The Transitional period has the highest score of `r trans_perm_r` while the Early Aurignacian has the lowest score of `r early_perm_r`. We see a large decline in scores between the Transitional and the Proto-Aurignacian periods, before a slight increase in between the Early and Evolved Aurignacian periods. The pseudo-F statistic quantifies the difference between groups, with the Transitional phase having the highest value of `r trans_perm_f`, and the Early Aurignacian having the lowest value of `r early_perm_f`. Similar to the R^2^ statistic, we see a decline in values between the Transitional and the Proto-Aurignacian periods, before a slight increase in between the Early and Evolved Aurignacian period. The perMANOVA p-value indicates a statistically significant difference in the distribution of signs between the restricted and broad groups for the Proto , Early, and Evolved Aurignacian phases, but not the Transitional phase (p = `r trans_perm_p`).

@@

## Grouping Dynamics Across the Four Phases of the Aurignacian

@@

## Social Organization Dynamics Throughout the Aurignacian

@@

@@

One limitation is the potential ambiguity that arises with some of the sign types. While @Von_Petzinger_2009 argues in favor of Aurignacian geometric signs as symbolic expression, potential ambiguity still remains over whether function is consistent across all sign type occurrences, particularly in regards to the simpler sign types composed from fewer elements. We included the more ambiguous sign types because similar sign occurrence may still serve as evidence of cultural connectivity regardless of function, and additionally, greater quantities of included sign types serves to establish stronger connections [@Gravel-Miguel_Coward_2023]. This ambiguity may be addressed in the future through the usage of the data recently released in SignBase 2.0 covering sites in Germany, which contains more detailed information on the specific order and patterns in which signs occurred on individual objects.
